---
title: "immdonor"
output: github_document
---
```{r global, include=FALSE}
library(flowCore)
library(flowWorkspace)
library(ggcyto)
library(openCyto)
library(MetaCyto)
library(cytoqc)
library(tidyverse)
library(knitr)
library(plotly)
library(kableExtra)
library(gtools)
```

Immdonor is a [GraphQL API](https://graphql.org/) of [Immport](https://www.immport.org/) Organ Donor metadata hosted on [Hasura Cloud](https://hasura.io/). This API lets you request exactly what data you need from the [Immport data model](https://www.immport.org/shared/dataModel) and returns a JSON output which can be converted to a R dataframe!

🚩This is a fast and easy way to link [Flow Cytometry Standard (.FCS)](https://en.wikipedia.org/wiki/Flow_Cytometry_Standard) files to their metadata for more robust analysis with [Cytoverse](https://cytoverse.org/). You will need to have the FCS files saved locally (see quick_fetch.R).

## Connect to API
```{r}
library(ghql)

#By connecting to this API you agree to the Immport.org Terms of Service
con <- GraphqlClient$new(url = "https://resolved-lab-57.hasura.app/v1/graphql")
```
***
#### GET
```{r}
qry <- Query$new()

studies<- qry$query('studies','{
  shared_data_study {
    study_accession
    actual_enrollment
    minimum_age
    maximum_age
  }
}')
x <- con$exec(qry$queries$studies)
```
#### ANSWER 
```{r, collapse= TRUE, warning=FALSE, message=FALSE}
studies<- as.data.frame (jsonlite::fromJSON(x))
```
```{r,echo=FALSE, warning=FALSE}
studies=studies[mixedorder(studies$data.shared_data_study.study_accession),]
studies=studies %>% select(data.shared_data_study.study_accession,data.shared_data_study.actual_enrollment,data.shared_data_study.minimum_age,data.shared_data_study.maximum_age)
row.names(studies) <- NULL

studies %>% kable(col.names = gsub("data.shared_data_study.", "", names(studies)))
```
***
## Organ Donor Metadata
The studies listed above feature cytometry data generously donated by organ donors. We can build queries to return exactly what metadata we would like to know about these organ donors.

```{r, collapse=TRUE, warning=FALSE, message=FALSE}
qry <- Query$new()

files<- qry$query('files','{
  resultFiles {
    filePath
    studyAccession
    subjectAccession
    maxSubjectAge
    gender
    biosampleType
    parameters
    panel_id
    markers
    
  }
}')

x <- con$exec(qry$queries$files)
files<- as.data.frame (jsonlite::fromJSON(x))
```

### Biosample types
```{r,echo=FALSE}
samples<- files %>% group_by(data.resultFiles.biosampleType) %>% count(data.resultFiles.biosampleType)

plo<- files %>% filter(data.resultFiles.biosampleType == "Thymus"|data.resultFiles.biosampleType == "Bone Marrow") %>% group_by(data.resultFiles.biosampleType) %>% count(data.resultFiles.biosampleType)

slo<- files %>% filter(data.resultFiles.biosampleType == "Spleen"|data.resultFiles.biosampleType == "Lung lymph node"|data.resultFiles.biosampleType == "Inguinal lymph node"|data.resultFiles.biosampleType == "Mesenteric lymph node"| data.resultFiles.biosampleType == "Tonsil") %>% group_by(data.resultFiles.biosampleType) %>% count(data.resultFiles.biosampleType)

mucosal<- files %>% filter(data.resultFiles.biosampleType == "Lung"|data.resultFiles.biosampleType == "Colon"|data.resultFiles.biosampleType == "Ileum"|data.resultFiles.biosampleType == "Jejunum"|data.resultFiles.biosampleType == "PBMC"|data.resultFiles.biosampleType == "Whole blood") %>% group_by(data.resultFiles.biosampleType) %>% count(data.resultFiles.biosampleType)

knitr::kables(
  list(
    knitr::kable(plo, col.names = c('Primary Lymphoid Organs', 'Count'), valign = 't' ),
    knitr::kable(slo, col.names = c('Secondary Lymphoid Organs', 'Count'), valign = 't'),
    knitr::kable(mucosal, col.names = c('Mucosal', 'Count'), valign = 't')
    ))


```
### Panels
Here are the top 5 staining panels with the most biosamples 
```{r,echo=FALSE}
panels<- files %>% group_by(data.resultFiles.panel_id,data.resultFiles.parameters) %>% count(data.resultFiles.markers)
panels=panels[order(panels$n, decreasing = TRUE),]
top10panels<- panels[1:5,]
top10panels %>% kable(col.names = gsub("data.resultFiles.", "", names(panels)))
```
***
## T Cell Analysis
```{r, echo=FALSE}
fcs<- files %>% filter(data.resultFiles.panel_id == "FCM-2"|data.resultFiles.panel_id == "FCM-3")

tpanels<- fcs %>% group_by(data.resultFiles.panel_id,data.resultFiles.parameters) %>% count(data.resultFiles.markers)
tpanels %>% kable(col.names = gsub("data.resultFiles.", "", names(tpanels)))

```

### Link cytoset to metadata
```{r, echo= TRUE, collapse= TRUE, message= FALSE, error= FALSE, eval=FALSE}
# Load cytoset 
fcs<- files %>% filter(data.resultFiles.panel_id == "FCM-2"|data.resultFiles.panel_id == "FCM-3")
fn<- fcs$data.resultFiles.filePath
cs<- load_cytoset_from_fcs(files = fn)

#link immdonor metadata to pData
p<- pData(cs)
m<- cbind(p,fcs)
pData(cs)<- m

# Harmonize marker names
channels <- colnames(cs)
markers <- as.vector(pData(parameters(cs[[1]]))$desc)
names(markers)<- channels
markernames(cs) <- markers

```

### Compensate and Transform
```{r, echo= TRUE, collapse= TRUE, message= FALSE, error= FALSE, eval=FALSE}
# Apply file internal compensation
comps <- lapply(cs, function(cf) spillover(cf)$SPILL)
cs_comp <- compensate(cs, comps)

# Transform fluorescent channels with FCSTrans
channels_to_exclude <- c(grep(colnames(cs), pattern="FSC"),
                         grep(colnames(cs), pattern="SSC"),
                         grep(colnames(cs), pattern="Time"))
chnls <- colnames(cs)[-channels_to_exclude]

fcstrans<- FCSTransTransform(transformationId = "defaultFCSTransTransform", channelrange = 2^18, channeldecade = 4.5, range = 4096, cutoff = -111, w = 0.5, rescale = TRUE)
transList <- transformList(chnls, fcstrans)

cs_trans<- transform(cs_comp,transList)

cs<- save_cytoset(cs_trans, "cytosets/tcell)
```

### Build Autogating strategy 
```{r, collapse= TRUE}
cs<- load_cytoset(path = "cytosets/tcell")
gs<- GatingSet(cs)

gs_add_gating_method(gs, alias = "nonDebris",
                     pop = "+",
                     parent = "root",
                     dims = "FSC-A",
                     gating_method = "mindensity",
                     gating_args = "min = 20000, max=50000",
                     collapseDataForGating = "TRUE",
                     groupBy = "data.resultFiles.biosampleType") 

gs_add_gating_method(gs, alias = "singlets",
                     pop = "+",
                     parent = "nonDebris",
                     dims = "FSC-A,FSC-H",
                     gating_method = "singletGate")

gs_add_gating_method(gs, alias = "CD19",
                     pop = "+/-",
                     parent = "singlets",
                     dims = "CD19",
                     gating_method = "mindensity",
                     gating_args = "min = 1700, max=3000",
                     collapseDataForGating = "TRUE",
                     groupBy = "data.resultFiles.biosampleType")


lln<- subset(gs, data.resultFiles.biosampleType == "Lung lymph node")
lung<- subset(gs, data.resultFiles.biosampleType == "Lung")
spleen<- subset(gs, data.resultFiles.biosampleType == "Spleen")


ggcyto(gs_pop_get_data(lln ,"CD19-"), aes(x = "CD4", y = "CD8")) + geom_hex(bins = 128)

```

* Coming soon:
  + Metacyto analysis
  + opencyto gating 
  + CyTOF data support


* Please contact me with ANY questions, comments or suggestions!!!
  + Slack: Join the [immunespace slack!](https://immunespace.slack.com/archives/DFU59KGUD)
  + Discord: MRJ#0408
  + email: mrjaffery at gmail dot com







