---
title: "immdonor"
output: github_document
---
```{r global, include=FALSE}
library(flowCore)
library(flowWorkspace)
library(ggcyto)
library(openCyto)
library(MetaCyto)
library(cytoqc)
library(tidyverse)
library(knitr)
library(plotly)
library(kableExtra)
library(gtools)

```

Immdonor is a [GraphQL API](https://graphql.org/) of [Immport](https://www.immport.org/) metadata hosted by [Hasura Cloud](https://hasura.io/). This API lets you request exactly what information you need from the [Immport data model](https://www.immport.org/shared/dataModel) and returns a JSON output which can be converted to a R dataframe. 

🚩This is a fast and easy way to link FCS files to their metadata and compliments the excellent [ImmportR](https://github.com/RGLab/ImmPortR) package. You still need to have the FCS files saved locally (see quick_fetch.R)

## Connect to Graphql API
```{r}
library(ghql)

#By connecting to this API you agree to the Immport.org Terms of Service
con <- GraphqlClient$new(url = "https://resolved-lab-57.hasura.app/v1/graphql")
```


***

### POST
```{r}
qry <- Query$new()

studies<- qry$query('studies','{
  shared_data_study {
    study_accession
    actual_enrollment
    minimum_age
    maximum_age
  }
}')

```
### Answer 
```{r, collapse= TRUE, warning=FALSE, message=FALSE}
x <- con$exec(qry$queries$studies)
studies<- as.data.frame (jsonlite::fromJSON(x))
```

```{r,echo=FALSE, warning=FALSE}
studies=studies[mixedorder(studies$data.shared_data_study.study_accession),]
studies=studies %>% select(data.shared_data_study.study_accession,data.shared_data_study.actual_enrollment,data.shared_data_study.minimum_age,data.shared_data_study.maximum_age)
row.names(studies) <- NULL

studies %>% kable(col.names = gsub("data.shared_data_study.", "", names(studies)))
```
***
## Organ Donor Metadata
The studies listed above feature cytometry data generously donated by organ donors. We can build queries to return exactly what metadata we would like to know about these organ donors.

```{r, collapse=TRUE, warning=FALSE, message=FALSE}
qry <- Query$new()

files<- qry$query('files','{
  resultFiles {
    filePath
    studyAccession
    subjectAccession
    maxSubjectAge
    gender
    biosampleType
    parameters
    panel_id
    markers
    
  }
}')

x <- con$exec(qry$queries$files)
files<- as.data.frame (jsonlite::fromJSON(x))
```

### Biosample types
```{r,echo=FALSE}
samples<- files %>% group_by(data.resultFiles.biosampleType) %>% count(data.resultFiles.biosampleType)

plo<- files %>% filter(data.resultFiles.biosampleType == "Thymus"|data.resultFiles.biosampleType == "Bone Marrow") %>% group_by(data.resultFiles.biosampleType) %>% count(data.resultFiles.biosampleType)

slo<- files %>% filter(data.resultFiles.biosampleType == "Spleen"|data.resultFiles.biosampleType == "Lung lymph node"|data.resultFiles.biosampleType == "Inguinal lymph node"|data.resultFiles.biosampleType == "Mesenteric lymph node"| data.resultFiles.biosampleType == "Tonsil") %>% group_by(data.resultFiles.biosampleType) %>% count(data.resultFiles.biosampleType)

mucosal<- files %>% filter(data.resultFiles.biosampleType == "Lung"|data.resultFiles.biosampleType == "Colon"|data.resultFiles.biosampleType == "Ileum"|data.resultFiles.biosampleType == "Jejunum"|data.resultFiles.biosampleType == "PBMC"|data.resultFiles.biosampleType == "Whole blood") %>% group_by(data.resultFiles.biosampleType) %>% count(data.resultFiles.biosampleType)


knitr::kable(
  list(plo, slo, mucosal),
  booktabs = TRUE ,align = "l" ,
  col.names = gsub("data.resultFiles.", "", names(samples))
)

```
### Panels
Here are the top 10 staining panels with the most biosamples 
```{r,echo=FALSE}
panels<- files %>% group_by(data.resultFiles.panel_id,data.resultFiles.parameters) %>% count(data.resultFiles.markers)
panels=panels[order(panels$n, decreasing = TRUE),]
top10panels<- panels[1:10,]
top10panels %>% kable(col.names = gsub("data.resultFiles.", "", names(panels)))
```


Coming soon:
-Metacyto analysis
-opencyto gating 
-CyTOF data support





